---
title: "DATA 607 Tidy Verse assignment"
author: "Yohannes Deboch"
date: "February 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 1.Create an Example.Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

## Load the library
```{r}
library(tidyverse)
```
## Data source
https://www.kaggle.com/ronitf/heart-disease-uci

## Read the data using readr

```{r}
disease <- read_csv("heart.csv")
head(disease)
```


The selected package I want to use id dplyr.

## Capability 1.

### filter capability tutorial

### Description

Using filter we can select rows of the data frame matching conditions.

### Usage
filter(data)
### Example

To select the people of over 20 and less than 65 we can pass the data disease and condtion age>20 and age < 65 to the function . It'll return matching rows of heart disease.

```{r}
filter(disease, age>20 & age < 65)
```

## Capability 2.

### select capability tutorial

###Description

Using select we can keep the selected variables

###sage
select(data, ...)

###Example

To keep only age, sex,cp variable we can pass the data disease and age, sex,cp to the function . 

```{r}
df<- select(disease, c("age","sex","cp"))
head(df)
```