---
title: "ds607_dplyr_tidyr_vignette"
author: "Jagdish Chhabria"
date: "April 28, 2019"

output: 
    html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_depth: 2
    toc_float: yes
---

## Introduction

In this vignette, I attempt to show the differences in how certain operations are performed using  base R versus using Tidyverse packages dplyr and tidyr. 

### Data

I will be using a dataset from Kaggle to demonstrate the differences in base R versus Tidyverse functions. This dataset provides information on 13 attributes of 1000 different camera models. It can be found at the following link; https://www.kaggle.com/crawford/1000-cameras-dataset/

Let's start by loading the data.

```{r}
# Load the required packages

library(readr)
library(tibble)
library(dplyr)
library(tidyr)

# Read in the required dataset and look at its column names and first few rows

cameras<-read.csv("https://raw.githubusercontent.com/Jagdish16/jagdish_r_repo/master/DATA607/camera_dataset.csv")

head(cameras)

colnames(cameras)
```

### Extracting variables

```{r }

# In base R

head(cameras[, c("Model", "Max.resolution", "Low.resolution","Effective.pixels", "Price")]) # by variable name
head(cameras[, c(1,3,4,5,13)])  # by column index


# In tidyverse

head(select(cameras, Model,Max.resolution, Low.resolution,Effective.pixels, Price)) # by variable name
head(select(cameras, 1,3,4,5,13))  # by column index

```


### Making new variables

```{r}

# In base R

cameras$Average.resolution = (cameras$Max.resolution + cameras$Low.resolution)/2
head(cameras$Average.resolution)


# In tidyverse

head(mutate(cameras, 
       Average.resolution = ((cameras$Max.resolution+cameras$Low.resolution)/2)
       ))

```


### Extracting observations

```{r}

# In base R

# Using subset
subset(cameras, Weight..inc..batteries. < 400 & Release.date == "2005")

# Using [,]
#cameras[Weight..inc..batteries. < 400 & Release.date == "2005", ]


# In tidyverse

filter(cameras, Weight..inc..batteries. < 400 & Release.date == "2005")


```


### Sorting observations

```{r}

# In base R

# descending order of price
head(cameras[order(rev(cameras$Price)),])

# In tidyverse

# descending order of species
head(arrange(cameras, desc(Price)))


```



### Summarizing observations

```{r}

# In base R

data.frame(Price.mean = mean(cameras$Price),
           Price.sd = sd(cameras$Price)
           )

# In tidyverse

summarise(cameras, 
          Petal.Length.mean = mean(Price),
          Petal.Length.sd = sd(Price),
        )

```

The above examples illustrate the differences in how certain common operations in data wrangling are carried out in tidyverse vis-a-vis base R. Hope this vignette helps demonstrate the difference in the 2 approaches.


