---
title: "CUNY MSDS Data 607 TidyVerse"
date: "April 26, 2019"
output: html_document
---

*** 

## { .tabset}

> Jimmy Ng's Tidyverse Recipe

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Code can be found on [Jimmy Ng's Github](https://github.com/myvioletrose/SPRING2019TIDYVERSE/blob/master/tidyverse_recipe_case_when_JN.Rmd).**

Let's load the tidyverse package and then load a sample data set from my github account. The original dataset comes from [kaggle](https://www.kaggle.com/vjchoudhary7/customer-segmentation-tutorial-in-python#Mall_Customers.csv). We will look at packages under the tidyverse, functions within "dplyr", and eventually zoom in into a specific dplyr function "case_when" with a sample use case.

***
### Set-up
```{r set-up, include = T, message = F, warning = F}
# load package(s)
library(tidyverse)
library(janitor)
library(kableExtra)
# load data
df <- read.csv("https://raw.githubusercontent.com/myvioletrose/school_of_professional_studies/master/607.%20Data%20Acquisition%20and%20Management/%23%20misc/Mall_Customers.csv") %>% 
        janitor::clean_names(.)  # using janitor::clean_names() to clean up column names
head(df) %>% kable() %>% kable_styling()
```

***
### tidyverse packages
```{r tidyverse_packages}
# how many packages live under the tidyverse?
tidyverse_packages()
```

***
### dplyr
```{r dplyr}
# how many functions within "dplyr"?
ls(pos = "package:dplyr")
```

***
### dplyr::case_when

Here's an official description: *This function allows you to vectorise multiple if_else() statements. It is an R equivalent of the SQL CASE WHEN statement. If no cases match, NA is returned.*
```{r case_when}
# let's create a new variable using case_when by transforming a numeric variable into buckets
# let's look at the distribution - before transformation
hist(df$annual_income_k, main = "Annual Income (before transformation)")
# let's create 10 buckets to represent this numeric variable
buckets <- cut(df$annual_income_k, breaks = 10) %>% levels
buckets
# create a new categorical variable
df$income_bucket <- cut(df$annual_income_k, breaks = 10)
df <- df %>%
        dplyr::mutate(income_bucket = dplyr::case_when(income_bucket == buckets[1] ~ "income level 0",
                                                       income_bucket == buckets[2] ~ "income level 1",
                                                       income_bucket == buckets[3] ~ "income level 2",
                                                       income_bucket == buckets[4] ~ "income level 3",
                                                       income_bucket == buckets[5] ~ "income level 4",
                                                       income_bucket == buckets[6] ~ "income level 5",
                                                       income_bucket == buckets[7] ~ "income level 6",
                                                       income_bucket == buckets[8] ~ "income level 7",
                                                       income_bucket == buckets[9] ~ "income level 8",
                                                       income_bucket == buckets[10] ~ "income level 9",
                                                       TRUE ~ as.character(income_bucket)))
head(df) %>% kable() %>% kable_styling()
# let's look at the distribution - after transformation
table(df$income_bucket) %>% barplot(., ylab = "Frequency", main = "Annual Income (after transformation)")
```


***

## { .tabset}

> Samantha Deokinanan's Extentions

I found [Jimmy Ng's](https://github.com/myvioletrose/SPRING2019TIDYVERSE/blob/master/tidyverse_recipe_case_when_JN.Rmd) tidyverse recipe very helpful as I have never used `dplyr::case_when`, thus I am extending this discussion with two (2) other `dplyr` functions, namely `dplyr::if_else` and `dplyr::recode`. I selected these fuctions because I thought they could have produced the same outputs as `case_when` (particularly `recode`) but there all have there differences and usefulness.   

***
### dplyr::if_else

This function is more strict than `base::ifelse` such that it will preserve the data type. `if_else` work great when a single set of conditions is to be satisfied. But if multiple sets of conditions are to be tested, nested `if_else` statements become cumbersome and thus prone to error. The following code highlights an example of nested `if_else` statements from what Jimmy Ng demonstrated:

```{r}
# Reset to df just before Jimmy Ng applied `dplyr::case_when` 
# i.e. create a new categorical variable
df$income_bucket <- cut(df$annual_income_k, breaks = 10)

df1 <- df %>%
        dplyr::mutate(income_bucket = 
                        dplyr::if_else(income_bucket == buckets[1], "income level 0",
                        dplyr::if_else(income_bucket == buckets[2], "income level 1",
                        dplyr::if_else(income_bucket == buckets[3], "income level 2",
                        dplyr::if_else(income_bucket == buckets[4], "income level 3",
                        dplyr::if_else(income_bucket == buckets[5], "income level 4",
                        dplyr::if_else(income_bucket == buckets[6], "income level 5",
                        dplyr::if_else(income_bucket == buckets[7], "income level 6",
                        dplyr::if_else(income_bucket == buckets[8], "income level 7",
                        dplyr::if_else(income_bucket == buckets[9], "income level 8", "income level 9"))))))))))

head(df1) %>% kable() %>% kable_styling()

str(df1)
```

Since `buckets` were already stored as character, the `income_buckets` categories were preserved without any errors. If `buckets` were numeric, then it would have to be converted into characters beforehand.

***
### dplyr::recode

Another solution is to use the `dplyr::recode` function, as it is well suited for replacing values but it will not allow for more complex operations. Let's look at it again with Jimmy Ng's example:

```{r}
df2 <- df %>%
        mutate(income_bucket = 
dplyr::recode(income_bucket, "(14.9,27.2]" = "income level 0",
                             "(27.2,39.4]" = "income level 1",
                             "(39.4,51.6]" = "income level 2",
                             "(51.6,63.8]" = "income level 3",
                             "(63.8,76]" = "income level 4",
                             "(76,88.2]" = "income level 5",
                             "(88.2,100]" = "income level 6",
                             "(100,113]" = "income level 7",
                             "(113,125]" = "income level 8", 
                             "(125,137]" = "income level 9"))

head(df2) %>% kable() %>% kable_styling()

str(df2)
```

Here, the recoded `income_bucket` was stored as a factor. 

***
### Conclusion

In the end, `dplyr` provides multiple functions that can be used when you want to create an output value that is conditioned on an evaluation. The right function can make the transformation less cumbersome.

***