---
title: 'TidyVerse: Purrr and Stringr/ ggplot2'
author: "Original by Javern Wilson - Extended by Mohamed"
date: "April 24, 2019"
output: 
  html_document:
    toc: false
    code_folding: show
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: flatly
    highlight: kate
---
## TidyVerse Task 2 - Extend an Existing Example {.tabset .tabset-fade .tabset-pills}

<li>I have extended <mark>Javern Wilson</mark> analysis to support some additional analysis with the help of graphical illustration using ggplot2 package</li> 
<li>Please navigate the tab below for orignial and extended version code</li>
<br>
Javern Wilson (Orginal) Code location: https://github.com/acatlin/SPRING2019TIDYVERSE/blob/master/DATA607%20Tidyverse%20Assignment.Rmd<br>
Mohamed (Extended) Code location : https://github.com/acatlin/SPRING2019TIDYVERSE/blob/master/tidyverse_task2_mohamed.Rmd
<br><br>

### By Javern Wilson (Orginal)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(kableExtra)
```

<h2>Purrr</h2>

Let's say we have a dataset of alcohol consumption among countries and we want to find the mean consumption of beer, spirit, wine and pure alcohol. Dataset was retrieved from [fivethirtyeight](https://github.com/fivethirtyeight/data/blob/master/alcohol-consumption/README.md).

```{r}
alcohol <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/alcohol-consumption/drinks.csv", sep = ",", stringsAsFactors = F)
alcohol
```

<br/>

With base R we would type:

```{r}
beer <- mean(alcohol$beer_servings)
spirit <- mean(alcohol$spirit_servings)
wine <- mean(alcohol$wine_servings)
pure <- mean(alcohol$total_litres_of_pure_alcohol)

c(beer, spirit, wine, pure)

```


As you can see, writing code like this is not very efficient because it involves a lot of copy and pasting and can make possible errors. To solve this issue of minimizing repetition with further replication, we can use `purrr`. `purrr` enhances R's functional programming (FP) toolkit by providing a complete and consistent set of tools for working with functions and vectors. 

`purrr` allows you to *map* functions to data. 

+ map() makes a list.
+ map_lgl() makes a logical vector.
+ map_int() makes an integer vector.
+ map_dbl() makes a double vector.
+ map_chr() makes a character vector.

Appropriately the basic function in purrr is called map() and it transforms the input by applying a function to each element and returning a vector the same length as the input.

We can use this function to perform the same computations as above.

```{r}
map_dbl(alcohol[, c(2, 3, 4, 5)], mean)

```

<br/> 

Pipes can also be used...
```{r}
alcohol[, c(2, 3, 4, 5)] %>% map_dbl(mean)
```

<br/>

#### Find other 'purrr' functions [here](https://github.com/rstudio/cheatsheets/blob/master/purrr.pdf)

<br/>

<h2> Stringr</h2>

This package is used to manipulate strings. For instance, if we want to gather the countries beginning with letter "S" using [regular expressions](https://www.rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf) or [regexs](https://medium.com/factory-mind/regex-tutorial-a-simple-cheatsheet-by-examples-649dc1c3f285). Patterns in stringr are interpreted as regexs.

```{r}
unlist(str_extract_all(alcohol$country, "^S.+"))

```


Or count how many countries begining with "S"
```{r}
str_count(alcohol$country, "^S.+")

#Sum up
sum(str_count(alcohol$country, "^S.+"))

```

<br/>

Another is to convert the names of the columns to a title format instead of doing so individually
```{r}
#original 
names(alcohol)
kable(head(alcohol)) %>% kable_styling()

#to title
names(alcohol) <- str_to_title(names(alcohol), locale = "en")
names(alcohol)
kable(head(alcohol)) %>% kable_styling()
```

<br/>

#### There are many other functions within the 'stringr' package that can be checked out [here](https://github.com/rstudio/cheatsheets/blob/master/strings.pdf)

### By Mohamed (Extended Analysis)

<h2>ggplot2</h2>

<li>Identified top 10 countries consuming high level of alcohol</li>

<li>Have used <mark>gather</mark> function from dplyr to transpose columns to rows</li>

<li>Drawn Stacked Bar Chart to Visualize the Type of Servings among those 10 countries</li>
<br>

```{r}
alcohol_Top10 <- alcohol %>% arrange(desc(Total_litres_of_pure_alcohol)) %>% top_n(10) %>% select(Country,Beer_servings,Spirit_servings,Wine_servings) 

alcohol_gather <- alcohol_Top10 %>% gather(Type, Servings,Beer_servings,Spirit_servings,Wine_servings) %>% arrange(desc(Servings,Country))

kable(alcohol_Top10) %>% kable_styling()

ggplot(data = alcohol_gather, aes(x = Country, y = Servings, fill = Type)) + geom_col() + coord_flip() + scale_fill_brewer(palette = 12) + labs(title = "Top 10 alcohol Consuming Countires - Servings Type")

```


