---
title: "Tidyverse Assignment"
author: "David Apolinar"
date: "5/3/2019"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("magrittr")
library("DT")
```

# Introduction

For This tidyverse assignment, I choose to use the following libraries:

* dplyr
* ggplot2
* magrittr
* forcats

I did not do extensive data filtering for this assignment, but rather choose to demostrate the capabilities of the tidyverse libraries.

# Data Pull

For my dataset, I choose to use kaggle with a gross avg rent across the US. The dataset is fairly long and has several fields. Not all are useful, but in the TidyVerse section, I use some of the ddplyr libaries to filter out what I need.

```{r rent data}
rent.data <- read.csv("https://raw.githubusercontent.com/dapolloxp/spring2019tidyverse/master/kaggle_gross_rent.csv")
```

# TidyVerse Libraries

In this section, I'm using some of the tidyverse libaries, such as piping in magrittr and dplyr which make it extremely easy to filter and select fields. I also chose fields using  select in ddplyr and group_by to aggregrate information.

I choose to count the number of zip codes with the piping built-into dplyr

```{r }
sub.rent.data <- rent.data %>% select(State_Name, State_ab, County, City, Place, Type, Primary, Zip_Code, Lat, Lon, Mean, Median, Stdev, Samples)

sub.rent.data$Zip_Code <- as.character(sub.rent.data$Zip_Code) 

head(sub.rent.data)
#datatable(sub.rent.data)

x <- function(zips)
{
  
  for (i in 1:length(zips)) 
  {
  paste0("Processing", zips[i])
  if(nchar(zips[i]) == 4)
    {
      zips[i] <- paste0("0", zips[i], sep = "")
    }
  
  else if(nchar(zips[i]) == 3)
    {
      zips[i] <- paste0("00", zips[i], sep ="")
    }
  }
  return(zips)
}
#sapply(sub.rent.data$Zip_Code, x)

#

sub.rent.data$Zip_Code <- as.factor(sub.rent.data$Zip_Code) 

sub.rent.data$Zip_Code %>% fct_count(sort = TRUE)


avg.rent.byzip <- sub.rent.data %>% group_by(Zip_Code) %>% summarise(AvgRent=mean(Mean))

avg.rent.bystate <- sub.rent.data %>% group_by(State_Name) %>% summarise(AvgRent=mean(Mean))
```

# Graphing

In this section, I am using ggplot to create a basic scatterplot as it is one of the most useful libraries.

```{r}
ggplot(avg.rent.byzip, aes(x=Zip_Code, y=AvgRent)) + geom_point()
ggplot(avg.rent.bystate, aes(x=avg.rent.bystate$State_Name,y=AvgRent)) + geom_point() + theme(axis.text.x = element_text(angle = 90)) + xlab("State Name") + ylab("Average Rent")
```