---
title: "Tidyverse Assignment"
author: "Mohamed Thasleem Kalikul Zaman"
date: "May 4, 2019"
output: 
  html_document:
    df_print: paged
    theme: united
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---
# Assignment

<li>In this assignment, I am using US hate crime data from fivethirtyeight website, compare and visualize the hate crime and how much share contributed on some factors like unemployment, with high school degree and non-citizen</li>   

# Data Used

<li>Data fetched from https://raw.githubusercontent.com/fivethirtyeight/data/master/hate-crimes/hate_crimes.csv</li>
<li>More information about data in https://github.com/fivethirtyeight/data/tree/master/hate-crimes</li>

# Packages Used

<li>usmap - US map graphical representation</li>
<li>ggplot2 - Utilized the capability of ggplot along with usmap package</li>
<li>dplyr - some data manipulation like selecting columns and mutating data values when there is 'NA' value</li><br>

```{r init_lib, warning=FALSE, message=FALSE}
library(tidyverse)
library(kableExtra)
library(usmap)
library(ggplot2)

hate_crimes <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/hate-crimes/hate_crimes.csv", sep = ",", stringsAsFactors = F)

per_100k_splc <- hate_crimes %>% select(state,avg_hatecrimes_per_100k_fbi,share_non_citizen,share_unemployed_seasonal,share_population_with_high_school_degree) %>% mutate(share_non_citizen = replace_na(share_non_citizen, 0)) %>% mutate(avg_hatecrimes_per_100k_fbi = replace_na(avg_hatecrimes_per_100k_fbi, 0))
                             
kable(data.frame(per_100k_splc)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  row_spec(0, bold = T, color = "white", background = "#ea7872") %>%
    scroll_box(width = "100%", height = "300px")
```

#Visualization

```{r init_lib1, warning=FALSE, message=FALSE}


us_crime <- plot_usmap(data = per_100k_splc, values = "avg_hatecrimes_per_100k_fbi", lines = "red",labels = TRUE) + 
  scale_fill_continuous(
    low = "white", high = "red", name = "US Hate Crimes"
  ) + theme(legend.position = "right", plot.title = element_text(hjust = 0.5)) + labs(title = "US Hate Crimes")

non_citizen <- plot_usmap(data = per_100k_splc, values = "share_non_citizen", lines = "red",labels = TRUE) + 
  scale_fill_continuous(
    low = "white", high = "red", name = "share_non_citizen"
  ) + theme(legend.position = "right", plot.title = element_text(hjust = 0.5)) + labs(title = "Non Citizen Share of hate crime")

unemployment <- plot_usmap(data = per_100k_splc, values = "share_unemployed_seasonal", lines = "red",labels = TRUE) + 
  scale_fill_continuous(
    low = "white", high = "red", name = "share_unemployed_seasonal"
  ) + theme(legend.position = "right", plot.title = element_text(hjust = 0.5)) + labs(title = "Unemployment Share of hate crime")

high_school_degree <- plot_usmap(data = per_100k_splc, values = "share_population_with_high_school_degree", lines = "red",labels = TRUE) + 
  scale_fill_continuous(
    low = "white", high = "red", name = "with School Degree"
  ) + theme(legend.position = "right", plot.title = element_text(hjust = 0.5)) + labs(title = "With School Degree Share of hate crime")

us_crime

non_citizen

unemployment

high_school_degree
```

 
